%\begin{center}
%  \textbf{ARGONNE NATIONAL LABORATORY}
%\end{center}

Argonne National Laboratory has three main facilities for computing
resources, namely the Argonne Leadership Computing Facility,
the Joint Laboratory for System Evaluation, and the Laboratory
Computing Resource Center.

\vspace{-1ex}
\paragraph{Argonne Leadership Computing Facility.}
Allocations are available through the INCITE, ALCC, and Director's Discretionary programs.
%
\textbf{Aurora} is Argonne's first exascale computer, built on a Cray EX platform with
Intel Data Center GPU Max Series GPUs and Intel XEON Max Series CPUs with HBM, connected
via HPE Slingshot fabric. Aurora supports machine learning, data science, and traditional
HPC modeling and simulation workloads at exascale.
\textbf{Polaris} is a 44-petaflop CPU/GPU hybrid system comprising 560 nodes, each with
4 NVIDIA A100 GPUs, fully integrated with a 200-petabyte parallel file system. Polaris
serves as a pre-exascale development platform for scaling codes toward Aurora.
The \textbf{ALCF AI Testbed} provides a suite of next-generation AI accelerators---Cerebras,
SambaNova, Graphcore, and Groq---for evaluating the usability and performance of
machine-learning-based HPC applications on specialized hardware.


%\vspace{-1ex}
%\paragraph{Other Compute Resources.}
Argonne provides additional computing through: the Joint Laboratory for System
Evaluation (JLSE), hosting cutting-edge platforms including Intel and NVIDIA development
GPUs and an FPGA testbed; the Laboratory Computing Resource Center (LCRC), with
the Improv CPU cluster (825 nodes, AMD EPYC 7713), Swing GPU cluster (8$\times$ A100 per node),
and Bebop (664 nodes); the CELS General Computing Environment (GCE), offering
2~PB NFS and 4~PB S3 object storage with Globus transfers; and the Rapid
Prototyping Laboratory (RPL), a robotics and software hub supporting autonomous discovery
workflows.
