Lawrence Livermore National Laboratory (LLNL) provides world-leading computing
resources through the Livermore Computing Center, including leadership-class
supercomputers, advanced data infrastructure, and specialized platforms that
support large-scale simulation, data science, and machine learning workflows.

\vspace{-1ex}
\paragraph{Livermore Computing Center.} Allocations are available through the Advanced
Simulation and Computing (ASC) program, the Institutional Computing (IC)
program, and Director’s Discretionary allocations.

\textbf{El Capitan} is LLNL’s first exascale supercomputer, based on the HPE Cray EX
platform and powered by AMD Instinct MI300A accelerated processing units
(APUs). El Capitan is designed to support large-scale multiphysics simulation,
AI-assisted modeling, and high-fidelity predictive science at exascale.

\textbf{Sierra} is a leadership-class hybrid CPU/GPU system comprising IBM POWER9
processors and NVIDIA V100 GPUs. Sierra supports large-scale HPC simulation,
machine learning, and data-intensive workflows, and serves as a critical
production platform for multiphysics and AI-enabled scientific applications.

\textbf{Tuolumne} is a next-generation advanced technology system designed to provide
pre-exascale capability and early access to emerging architectures. Tuolumne
supports code development, performance portability testing, and scaling studies
for next-generation HPC systems.

\vspace{-1ex}
\paragraph{Other Compute Resources.} LLNL provides additional computing capabilities through:
Institutional Computing (IC) clusters that support moderate- to large-scale
simulation workloads; high-performance storage systems providing multi-petabyte
parallel file systems for large-scale simulation data; advanced visualization
and data analysis environments for large-scale scientific workflows; and
specialized AI and machine learning platforms that enable development and
evaluation of data-driven and hybrid simulation workflows.
