\textbf{Decision Gate Metrics and Go/No-Go Criteria}

Phase~I success will be evaluated using quantitative metrics that assess whether the proposed foundation-model and agentic-AI workflow provides a clear AI advantage over conventional CFD-only design. The decision gates focus on four areas: representation fidelity, predictive accuracy, reduction in expensive high-fidelity simulations, and end-to-end design acceleration. These metrics are aligned with the stated Phase~I objectives and milestones of the proposal.

{\bf 1. Latent representation fidelity.}
A first success criterion is that heterogeneous datasets from Nek5000/RS, MFEM, and experiments can be mapped into a common geometry-aware latent space without significant loss of information. Success will be defined by relative reconstruction errors below 3\% for the primary thermal-fluid state variables and below 1.5\% for key derived quantities such as pressure drop and heat-transfer metrics.

{\bf 2. Predictive accuracy on unseen cases.}
The foundation model must generalize beyond the training set. Success will be defined by mean prediction errors below 3\% for quantities of interest such as pressure drop, Nusselt number, heat-transfer coefficient, and selected mixing metrics on held-out geometries and operating conditions; and below 5\% for first- and second-order turbulence statistics. This level of accuracy must be demonstrated across at least three representative configuration classes.

{\bf 3. Reduction in required CFD evaluations.}
The AI-guided workflow must reduce reliance on repeated high-fidelity simulations during design. Success will be defined by demonstrating that the latent-space optimization identifies designs of comparable or better quality than a conventional parameter sweep or baseline optimization workflow while using at least 5$\times$ fewer high-fidelity CFD evaluations, with 10$\times$ as the target value.

{\bf 4. End-to-end design acceleration.}
The combined surrogate-plus-agentic workflow must reduce total design turnaround time relative to a CFD-only workflow. Success will be defined by demonstrating at least 10$\times$ acceleration in end-to-end design-cycle time on a representative heat-transfer or reactor thermal-hydraulics problem, with 20$\times$ as the stretch target. Additional evidence of AI advantage will be provided if increased training data and/or compute systematically improve either prediction accuracy or optimization efficiency.

{\bf Go/No-Go Criterion.}
Advancement to Phase II will require simultaneous demonstration of:
(i) prediction errors below 5\% on unseen cases;
(ii) at least 5$\times$ reduction in the number of required high-fidelity CFD evaluations; and
(iii) at least 10$\times$ reduction in end-to-end design-cycle time for a representative DOE-relevant design problem. Meeting these criteria will establish that the proposed framework delivers a credible and measurable AI advantage for accelerated thermal-fluid design.


\begin{comment}


Phase~I success will be evaluated through quantitative metrics
demonstrating predictive accuracy, computational efficiency,
and measurable AI advantage relative to conventional
high-fidelity CFD workflows using Nek5000/RS and MFEM.

\textbf{1. Integrated Data and Latent Representation.}
A unified geometry-aware dataset combining simulation
and experimental thermal–fluid data will be constructed.
Success criteria include:
%
Successful integration of $\geq 95\%$ of selected datasets
into a consistent latent representation;
%
Reconstruction fidelity of encoded flow fields with
relative $L_2$ error $\leq 3\%$ for representative
thermal–fluid variables.

\textbf{2. Predictive Accuracy and Generalization.}
The foundation model will be validated across unseen
geometries and operating regimes. Success requires:
%
Prediction error $\leq 5\%$ for engineering quantities
including pressure drop and heat-transfer coefficients;
%
Robust performance across at least three geometry
classes and multiple Reynolds-number regimes.

\textbf{3. Design Acceleration and Optimization Efficiency.}
AI-enabled latent-space optimization will be compared
to traditional brute-force parameter sweeps. Success
criteria include:
%
Reduction of required high-fidelity CFD evaluations
by $\geq 5$--$10\times$;
%
Demonstration of stable multi-objective optimization
balancing thermal performance and pressure-drop penalties.

\textbf{4. Demonstration of AI Advantage Through Scaling.}
AI advantage will be quantified using scaling-based
performance metrics:
%
Reduction in prediction error of $\geq 30\%$
when training dataset size is increased by $2\times$;
%
Efficient GPU scaling demonstrating proportional
reduction in training or inference time with increasing
computational resources;
%
Demonstration of surrogate inference throughput
$\geq 100\times$ faster than equivalent CFD evaluation.

\textbf{Go/No-Go Criterion.}
Advancement to Phase~II requires simultaneous
demonstration of:
Prediction accuracy $\leq 5\%$ across unseen geometries;
$\geq 5\times$ reduction in required CFD evaluations;
Measurable scaling-based AI advantage;
$\geq 10\times$ reduction in end-to-end
design-cycle time relative to traditional workflows.

Meeting these criteria will validate the feasibility of
foundation-model–driven workflows for accelerated
thermal–fluid design and justify Phase~II expansion.

\end{comment}
