\section{Milestones in the Nine Months}
\noindent
\textbf{Milestone 1 (Month 2): Data Integration and Baseline Surrogate}

Unify \textbf{Nek5000/RS} simulation and experimental datasets
(wire-coil and reactor-relevant geometries) into a common representation
and demonstrate a baseline surrogate model for heat-transfer and
pressure-drop prediction. 
Preparation of data sources and initial ICF simulations in MFEM.

\vspace{4pt}

\noindent
\textbf{Milestone 2 (Month 4): Foundation Model in Latent Space}

Fine-tune an existing geometry-aware fluid-mechanics foundation model
using project-specific datasets. The learned latent representations will
enable cross-configuration generalization, achieving accurate predictions
($<5\%$ error) across unseen geometries. 
Initial version of the MFEM-based ICF miniapp.
Continued collection of data for fuel-pin and wire-coil models.

\vspace{4pt}

\noindent
\textbf{Milestone 3 (Month 6): Closed-Loop Optimization Capability}

Integrate Bayesian and derivative-free optimization methods in latent
space and demonstrate closed-loop design loops, achieving at least
$5$--$10\times$ reduction in design iteration cost relative to
brute-force CFD in both Nek and MFEM based applications.

\vspace{4pt}

\noindent
\textbf{Milestone 4 (Month 9): Agentic AI--Driven Accelerated Design}

Deploy agentic AI orchestration for adaptive sampling and
multi-objective optimization, demonstrating end-to-end automated
design with $10$--$20\times$ design-cycle acceleration and improved performance
(e.g., heat-transfer versus pressure-drop trade-off) on a
representative heat-exchanger system and ICF configuration.
