\input figfm1

\section{Background/Introduction}

This project will develop a {\bf foundation-model-driven, agentic-AI framework}
\cite{Vinuesa2026XAI} for the prediction, optimization, and 
{\color{red}VINUESA: remove? control} of
thermal-fluid transport, with primary emphasis on components in fission and
fusion energy systems.
  Phase I demonstrator problems, illustrated in Fig.~\ref{fig:fm1}, right, 
include 
{\em (i)} wire-wrapped fuel-rod bundles, 
{\em (ii)} an inertial-confinement fusion (ICF) miniapp, and 
{\em (iii)} heat-transfer enhancement through optimized wire-coil-inserts.
  These components exemplify systems central to DOE’s advanced energy
programs and are designed to permit ready exploration, development, and
validation of agentic workflows in this space, where
    energy transfer is characterized by complex turbulent flow, strong
geometric sensitivity, and stringent performance and safety constraints.
  They also constitute a rich data domain, combining decades of high-fidelity
simulations using Nek5000/RS (one of the few CFD codes accepted in NRC
licensing workflows and a former Gordon Bell Prize winner
\cite{tufo99a,gb23}) and MFEM (the 2025 Gordon Bell Winner \cite{gb25mfem}) 
with extensive experimental datasets from both open literature and proprietary
industry sources.  These two teams have worked closely together in the last
decade to deliver major application and HPC advancements to the DOE under the
Exascale Computing Project \cite{ceed}.

%  The central objective of this {\bf Phase I effort} is to {\em unify the
%  available heterogeneous datasets into a common (mesh-based) geometry-aware
%  representation and to demonstrate rapid, accurate, surrogate solution}
%  generation that will be applied to design and optimization of thermal-fluid
%  systems.




Thermal–fluid transport governs the performance of a wide range of energy systems.
In fission and fusion systems, fluid motion transports thermal energy
from core processes (fission/fusion reactions) to heat exchangers, where 
recipient working fluids ultimately drive turbines or other power-generation devices.
Combustion systems and inertially confined fusion (ICF) platforms similarly
depend on controlled fluid motion to initiate and sustain energy-producing
reactions.
However, {\bf system- and device-level optimization—particularly when involving
geometric or topology modifications—typically requires tens to hundreds of
high-fidelity simulations}, each taking days to months on high-performance
computing (HPC) platforms,
with {\bf no shared information across the simulation space.} The computational
burden arises from complex geometries, broad ranges of timescales (e.g.,
convective flow versus thermal diffusion in solids), and stringent accuracy
requirements.

  The Phase I effort will combine graph neural networks (GNNs) with
autoencoders to build a cross-device latent space comprising compressed
geometric, fluid, and thermal information across {\em all} project data sets.  
This shared information space will permit significantly shortened
training cycles for {\em new device configurations} (e.g., fuel rods vs. heat
exchangers).  The quality of the surrogate predictions---with 
emphasis on design-driven quantities of interest (QOIs)---will be
assessed directly through experimental data and current simulation
capabilities.  The ultimate goal, to be realized in Phase II,
is to radically transform the design cycle for energy-system components by
providing reliable, high-accuracy AI-driven surrogate models tightly coupled
with state-of-the-art design optimization methods.  Runtime surrogate
uncertainty quantification (UQ) will be a key component of Phase II so that
necessary retraining may be identified at low cost.

% The Phase I test problems are selected to establish feasibility across
% representative DOE-relevant applications, including reactor-core flows and
% inertially confined fusion (ICF) miniapps, while liquid-metal
% magnetohydrodynamics (MHD) relevant to magnetically confined fusion will be
% addressed in Phase II.
% As illustrated in Fig. 1, Phase I test cases include:
%     (i) optimization of pressure-drop and heat-transfer performance in \textbf{wire-wrapped fuel-rod bundles};
%    (ii) a \textbf{miniapp to investigate approaches for inertially confined fusion}; and
%   (iii) optimization of \textbf{wire-coil inserts} for
% heat-transfer enhancement in beam-line cooling systems.
% These test cases were selected because of the availability of reliable simulation
% and experimental data and their scalability across a wide range of problem sizes,
% with simulation degrees of freedom (DOFs) ranging from $10^7$ to $10^{12}$
% \cite{gb23,gb25mfem}.
% {\bf PF: need to clean up / re-org; eliminate redundancy in the preceding paragraph.}


\section{Project Objectives}



\input objective
\noindent

\vspace*{.1in}
\section{Proposed Research and Methods}

% \tiny 
% \begin{verbatim} 
%   Provide a clear research plan for a 9-month project and describe the proposed activities and methods.  
%   Include enough technical details to evaluate the impact of the proposed activity.  
%   For each activity, indicate the responsibility of the key investigator(s) and the associated budget.  
%  *** If the proposed application is building on an existing, currently funded DOE 
%      project, describe how the submitted leverages that work and is distinct from 
%     the existing funding. ***  [MENTION ASCR/NEAMS ! ]
% \end{verbatim} 
% \normalsize


Figure~\ref{fig:tax} summarizes our approaches.
As illustrated in Fig.~\ref{fig:fm1}, our goal of {\bf Phase I} is to develop a training
workflow to build AI surrogates for thermal-fluid solutions. 
Initial training sets will be based on time-independent {mean-flow} fields that are already
available or readily generated from Nek5000/RS and MFEM using direct numerical
or large-eddy simulations (DNS or LES).  The geometry for each dataset is embedded
in the computational mesh, whose nodes are associated with corresponding
solution variables velocity, pressure, and temperature.   Cases with
varying geometry or parameters (e.g., Reynolds number) are fed into an
encoder, which builds the latent space. In this setting we will also be able to
take advantage of the ASCR-supported development of differential simulation
capabilities in MFEM in order to produce higher-quality surrogates or to
converge the optimization at high resolution given a surrogate initial guess
\cite{MFEM24, DMC24, Schmidt26}.    
%
This work leverages current DOE investments in high-fidelity simulation
through DOE's Nuclear Energy Advanced Modeling and Simulation (NEAMS) program
(e.g., \cite{smr21,Merzari2024Exascale} {\bf [more Dillon]}) and ASCR,
where in addition to reactor simulation capabilities, we are developing
magnetohydrodynamics (MHD) capabilities for simulating liquid-metal divertors
and breeder-blankets for magnetically-confined fusion {\bf [misun1,misun2,tzanio1]}.
The latter will be a critical component of Phase II developments.

Once trained, a new geometry (mesh) can be fed into the encoder to generate a
surrogate solution within a few seconds.  This process has been successfully
applied to LES-based flow data for the NASA full-aircraft test model at
different angles of attack and slat positions using datasets of size $n \approx
50$M.  The encoding process uses PyTorch on GPUs (several days on four to eight
H200s) and inference takes about 30 seconds.  As products of DOE's Exascale
Computing Project, both Nek5000/RS and MFEM run on NVIDIA, AMD, and Intel
accelerators, so there are no GPU portability issues.

This effort builds directly on DOE investments in high-fidelity simulation.
Nek5000/RS and MFEM developed under the Center for Efficient Exascale
Discretizations as part of DOE's Exascale Computing Project \cite{ceed}.  
Their high-order
discretizations, coupled with fast matrix-free operator evaluation and highly
scalable preconditioners gives them a competitive training edge over
traditional approaches (e.g., a 32$\times$ gain compared to 2nd-order finite volume
methods \cite{Min2024Wind,Tomboulides2024ABL}).  Nek5000/RS has been
extensively used for reactor thermal-hydraulics and is part of NRC-accepted
workflows, while MFEM has enabled important fusion applications in the DOE,
based both on magnetic and on inertial confinement approaches.
The proposed work, however, is fundamentally distinct.
Instead of advancing simulation capabilities alone, we develop a generalizable
foundation model that unifies data across geometries and regimes, and we
introduce a closed-loop, AI-driven optimization framework that enables rapid
design exploration and mechanistic discovery.   This process has already
been demonstrated for full aircraft configurations (co-PI Vinuesa).
By the end of Phase I, we will
demonstrate real-time surrogate modeling and surrogate-based optimization for
representative thermal-fluid systems, establishing a clear pathway toward
closed-loop optimization and AI-driven digital twins for reactor design 
and other DOE mission applications in Phase II.

\input figtax

The first component of the research plan, led by ANL, UIUC, and PSU, focuses on
the development of a {\bf geometry-aware thermal–fluid foundation model}.
Building on prior work in foundation modeling \cite{Vishwasrao2026AgenticPDE},
simulation and experimental data from wire-wrapped
bundles~\cite{Dai2025WireWrapped61Pin,kaeri37pin}, as well as related
configurations such as spacer grids, pebble beds, and twisted tubes, will be
mapped into {\rm a common latent space independent of mesh topology}. This
capability is particularly critical for reactor applications, where variations
in geometry—such as wire pitch, rod spacing, and bundle configuration—strongly
influence heat transfer and pressure losses.  The proposed model integrates
{\rm graph-based geometric representations}, {\rm variational autoencoders
(VAEs)} for compression of high-dimensional flow
fields~\cite{Vinuesa2022MLCFD,SoleraRico2024BetaVAE}, and {\rm conditional
latent diffusion models}~\cite{Vishwasrao2025DiffSPORT} to learn distributions
across operating regimes, as well as neural-operator architectures for
physics-consistent surrogate learning which we will investigate in Phase II.
Derivatives of the velocity and temperature fields are central to quantities of
interest (QOIs) such as pressure drop and heat transfer.  We will compute these
in the fine-scale training data, where derivatives can be accurately computed,
and encode them along with the primary data.  The resulting surrogate model
will enable fast, high-fidelity prediction of key reactor performance
quantities, including heat transfer coefficients, pressure drop, and
flow-induced mixing. For the cases considered, we have already established
clear agreement between the experimental data and the high-order simulation
codes, which will be used to validate the ML surrogates.

The MFEM team will contribute a {\bf targeted ICF miniapp} designed as a
high-fidelity data generator for training and validating the proposed surrogate
modeling framework. Building on existing capabilities in Lagrangian
hydrodynamics \cite{Dobrev2012}, ALE methods \cite{Tomov2018}, and scalable
H(div) radiation diffusion solvers \cite{Pazner2025}, this effort will focus on
producing physically consistent, multi-physics datasets that capture coupled
hydrodynamics–radiation behavior across parameterized geometries and operating
conditions. The miniapp will be tightly integrated into the AI workflow to
systematically generate training data, support active learning loops, and
enable evaluation of surrogate accuracy and generalization. In Phase I,
simplified ICF configurations will be used to demonstrate feasibility and
establish a robust pipeline connecting MFEM-generated data with the foundation
models, ensuring that learned representations remain consistent with the
underlying physical operators and constraints.
This effort directly complements the proposed AI architecture by providing
high-order, structure- preserving finite element data that aligns with the
project’s emphasis on physically consistent latent spaces. MFEM’s support for
complex, unstructured meshes and high-order discretizations enables systematic
variation of geometry and resolution, which is critical for training
geometry-aware models such as GNN-based encoders. Moreover, the compatibility
of MFEM operators with differentiable programming frameworks enables a natural
pathway to incorporate operator-level sensitivities into both surrogate
training and downstream optimization. This will create a strong link between the
simulation stack and the latent-space optimization strategies outlined in the
proposal, particularly for adjoint-informed refinement and hybrid optimization
workflows.

The second component, led by the University of Michigan, develops a {\bf
latent-space optimization and control framework}, targeting both steady and
transient operating regimes.  This effort will employ a hybrid strategy
combining {\rm Bayesian optimization}~\cite{Morita2022BOCFD}, {\rm
derivative-free optimization methods}, and 
{\color{red}VINUESA, remove? {\rm reinforcement learning}~\cite{Font2025DRLFlowControl}}.  
Such approaches are essential for
reactor-relevant design spaces that are high-dimensional, non-convex, and
constrained by safety and operational limits. Note that the MFEM team also
has experience with {\color{red}VINUESA: remove? reinforcement learning}
in for adaptive DOE-relevant
finite element simulations \cite{RL23}.
%
Optimization will be performed directly within the latent space of the
foundation model, enabling efficient exploration of design variables such as
wire pitch, rod arrangement, and operating conditions. The {\rm agentic AI
system} will orchestrate this process by dynamically selecting optimization
strategies, proposing new simulation queries, and balancing competing
objectives, including heat transfer enhancement, pressure loss minimization,
and safety margin preservation.
%
The framework will operate in a closed-loop workflow, in which candidate
designs generated in latent space are selectively validated using high-fidelity
simulations.  This ensures predictive accuracy while maintaining robustness
across operating conditions.
%
All team members will work on {\bf cross-domain validation and integration into the
Genesis Mission ecosystem}.



\vspace*{.1in}
\noindent {\bf Effort Breakdown:}
   {\bf ANL}-reactor TF data; model validation; device optimization \$100K;
   {\bf LLNL}-ICF data; model validation \$100K;
   {\bf U. Mich.}-model development; device optimization \$100K;
   {\bf PSU}-reactor TF data; model validation \$50K;
   {\bf UIUC}-heat-exchanger data; model validation \$50K.
{\tiny \bf CHANGE to REFLECT FULL BUDGETs}

\input milestone

\input data

\input metric
