The objective of this project is to develop a geometry-aware, foundation-model–driven,
agentic AI framework for the rapid prediction, optimization, and control of
thermal–fluid transport in DOE-relevant energy systems. The proposed work
directly aligns with the chosen focus topic on Artificial Intelligence–enabled
design and control of complex energy systems by integrating high-fidelity
simulation data, advanced machine learning architectures, and closed-loop
optimization workflows.
%
Specifically, the project will:
\begin{enumerate}
\item \textbf{Develop a unified data representation and baseline surrogate}
that integrates heterogeneous simulation and experimental datasets from
\textbf{Nek5000/RS} and \textbf{MFEM} into a common geometry-aware latent
space suitable for large-scale foundation-model training.

\item \textbf{Construct and fine-tune a geometry-aware foundation model}
that enables cross-configuration prediction of thermal–fluid quantities
(e.g., heat-transfer coefficients and pressure drop) with target accuracy
better than $5\%$ across unseen geometries.

\item \textbf{Enable latent-space optimization and closed-loop design}
through integration of Bayesian optimization, derivative-free methods,
and reinforcement learning, reducing the number of required high-fidelity
simulations by at least $5$--$10\times$ relative to brute-force CFD-based
design exploration.

\item \textbf{Deploy an agentic AI orchestration framework} that performs
adaptive sampling and multi-objective optimization, demonstrating
end-to-end automated design workflows with $10$--$20\times$ acceleration
in design-cycle time for representative heat-transfer systems.
\end{enumerate}

These objectives directly support the selected focus topic by enabling
AI-driven predictive modeling and automated design of complex thermal–fluid
systems central to advanced fission, fusion, and high-performance energy
technologies. The resulting AI advantage arises from the ability to replace
large ensembles of computationally expensive simulations with learned,
physics-informed surrogate models that generalize across geometries and
operating conditions. This capability reduces design-cycle cost, accelerates
innovation timelines, and enables exploration of previously inaccessible
design spaces.

%\textbf{Project Objectives and AI Advantage}
%
%The objective of this project is to develop a geometry-aware,
%foundation-model–driven, agentic AI framework for rapid prediction,
%optimization, and control of thermal–fluid transport in DOE-relevant
%energy systems. The work aligns with the selected focus topic by
%enabling AI-driven predictive modeling and automated design of
%complex energy components using high-fidelity simulation data
%from Nek5000/RS and MFEM.
%
%Key objectives include: (i) development of a unified geometry-aware
%latent representation integrating heterogeneous datasets;
%(ii) construction of a foundation model achieving prediction
%accuracy better than $5\%$ across unseen configurations;
%(iii) implementation of latent-space optimization achieving
%$5$--$10\times$ reduction in required high-fidelity simulations;
%and (iv) deployment of agentic AI workflows achieving
%$10$--$20\times$ acceleration in design-cycle time.
%
%The resulting AI advantage lies in replacing computationally
%expensive simulation campaigns with transferable,
%physics-informed surrogate models that enable rapid exploration
%of large design spaces while maintaining engineering-level accuracy.
