\subsection{Ricardo, AI for fluids}

This project will develop a foundation-model-driven, agentic-AI framework~\cite{Vinuesa2026XAI} for the prediction, optimization, and control of thermal-fluid transport in energy systems, with primary emphasis on wire-wrapped fuel-rod bundles for advanced reactor design. These systems represent a cornerstone of DOE’s mission in advanced fission and fusion energy, and are characterized by complex turbulent flows, strong geometry dependence, and stringent performance and safety constraints. They also constitute a uniquely rich data domain, combining decades of high-fidelity simulations using Nek5000/RS (one of the few CFD codes accepted in NRC licensing workflows) with extensive experimental datasets from both open literature and proprietary industry sources. The central objective of this Phase I effort is to unify these heterogeneous datasets into a common representation and demonstrate a closed-loop workflow for rapid design and optimization.

The first component of the research plan focuses on the development of a geometry-aware foundation model for thermal-fluid prediction, led by Argonne National Laboratory. Building on prior work on foundation models~\cite{Vishwasrao2026AgenticPDE}, as well as simulation and experimental data from wire-wrapped bundles~\cite{Collins2022WireCoil} and related configurations such as spacer grids, pebble beds, and twisted tubes, will be mapped to a common latent space independent of mesh topology. This is particularly critical for reactor applications, where geometry variations (wire pitch, rod spacing, bundle configuration) strongly influence heat transfer and pressure losses. The model will combine graph-based representations of geometry, variational autoencoders for compressing high-dimensional flow fields~\cite{Vinuesa2022MLCFD,SoleraRico2024BetaVAE}, and conditional latent diffusion~\cite{Vishwasrao2025DiffSPORT} to learn the latent distribution across operating regimes. Training will exploit DOE exascale-ready workflows and GPU-enabled datasets from Nek5000/RS and MFEM. The resulting surrogate will provide fast, accurate predictions of quantities central to reactor design and licensing, including heat transfer coefficients, pressure drop, and flow-induced mixing.

The second component develops a latent-space optimization and control framework, led by the University of Michigan, targeting both steady and transient regimes. We will employ a hybrid strategy combining Bayesian optimization~\cite{Morita2022BOCFD}, derivative-free methods, and reinforcement learning~\cite{Font2025DRLFlowControl}. This is essential for reactor-relevant design spaces, which are high-dimensional, non-convex, and constrained by safety and operational limits. Optimization will be performed directly in the latent space of the foundation model, enabling efficient exploration of design parameters such as wire pitch, rod arrangement, and operating conditions. The agentic-AI system will orchestrate this process by selecting optimization strategies, proposing new simulation queries, and balancing competing objectives such as heat transfer enhancement, pressure loss, and safety margins. The framework will operate in a closed loop, where candidate designs generated in latent space are selectively validated using high-fidelity simulations, ensuring both accuracy and robustness.

The third component focuses on validation across multiple DOE-relevant applications and integration into the Genesis Mission ecosystem, led by Argonne with contributions from all partners. In addition to wire-wrapped reactor bundles, we will consider two complementary applications that expand the scope and generality of the approach. The first is inertial confinement fusion, where highly transient, multiphysics dynamics arise from laser-driven compression of fuel pellets. This provides a stringent test of the model’s ability to capture fast-evolving regimes. The second is a well-characterized heat transfer enhancement problem involving wire-coil inserts in cooling channels, where experimental data show up to fourfold increases in heat transfer with well-quantified pressure penalties . This problem is particularly valuable for model development and validation due to its rich parameter space (wire diameter, pitch, Reynolds number) and the strong agreement between simulations and experiments. Together, these applications ensure both depth (reactor relevance) and breadth (generalization across physics and regimes).

This effort builds directly on DOE investments in high-fidelity simulation through the Exascale Computing Project, particularly Nek5000/RS, which has been extensively used for reactor thermal-hydraulics and is part of NRC-accepted workflows. However, the proposed work is fundamentally distinct. Instead of advancing simulation capabilities alone, we develop a generalizable foundation model that unifies data across geometries and regimes, and we introduce a closed-loop, AI-driven optimization framework that enables rapid design exploration and mechanistic discovery. By the end of Phase I, we will demonstrate real-time surrogate modeling and closed-loop optimization for representative thermal-fluid systems, establishing a clear pathway toward AI-driven digital twins for reactor design and other DOE mission applications.