\subsection{Sven AI, Digitial Twins}

We will develop a closed-loop differentiable agentic franmework for the optimization tasks arising in the design and control of nuclear reactors. Our framework builds on surrogate models trained on both highh-fidelity simulations from MFEM or NEK and on adjoints available from MFEM. This co-design of surrogates and adjojnts ensures that the surrogate has a gradient that is consistent with the gradient of the simulation. This approach is referred to a derivative-informed neural operator (DINO), see \cite{o2024derivative,luo2025efficient,cao2025derivative,wu2023large}.

In general, DINO is a family of neural networks that approximate infinite-dimensional mappings such as solutions to PDEs. DINO co-trains booth function and adjoint information, exploiting the intrinsic low-dimensionality of the derivatives by compressing derivative-information and imposing this information during the neural operator training. DINO can be trained at a marginal additional cost compared to traditional operators without consistent gradient information and achieves higher solution accuracy than traditional methods. DINO have been applied to solving PDEs with high-dimensional uncertain parameter space that would otherwise have been intractable.

The availability of consistent gradient information allows us to apply modern gradient-based optimization techniques to the optimal design and control problems in this project. In general, we cannot expect the neural surrogates to perfectly match the high-fidelity simulations at the optimal design point. Hence, we propose to embed the optimization solver within a differentiable agentic framework that leverages DINO and orchestrates high-fidelity simulations to augment or retrain the neural surrogate, and drive the design or control towards optimal parameters. In Phase II we will exploit the multilevel characteristics of our design and control problems within the agentic framework by allowing it to control the fidelity of our simulations. We envisage an outer-loop agentic framework kthat drives both the optimization and the fidelity of additional  simulations orchestrating the gradient-based optimization that employ DINO and the high-fidelity simutations from NEK and MFEM.

